---
title: "STA238 Final Project"
author: "Qiaoyu Wang & Zhanfei Gu"
date: '2022-03-21'
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(latex2exp)
```

## Data cleaning

```{r}
# Import data 
data <- read_csv("Life Expectancy Data.csv", show_col_types = FALSE)

# Data cleaning
# Remove missing value on variable Life expectancy, Total expenditure
# Only keep the data for year 2014
# Only keep the 5 variables in the final dataset

data_clean <- data %>%
  filter(!is.na(`Life expectancy`)& !is.na(`Total expenditure`)& Year == 2014) %>%
  select(Country, Year, Status, `Life expectancy`, `Total expenditure`)

```

## Exploratory Data Analysis

## Numerical Summary


```{r}
## units (age)
summary(data_clean$`Life expectancy`)
```
```{r}
## units (%)
summary(data_clean$`Total expenditure`)
```

## Graphical summary

```{r}
## Create a side-by-side boxplot to check the expenditure of developing and developed countries
ggplot(data = data_clean, aes(x = Status, y = `Total expenditure`)) +
  geom_boxplot(fill = "steelblue", color = "black") +
  labs(x = 'Country Status (Developed or Developing)', 
       y = 'Expenditure on health of total government expenditure(%)',
       title = "Box plot of total expenditure for developing and developed countries")

```

```{r}
## Create a histogram to see the distribution of life expectancy
ggplot(data = data_clean, aes(x = `Life expectancy`)) +
  geom_histogram(fill = 'steelblue', color = 'black', bins = 15) +
  labs(x = 'Life Expectancy (age)', title = 'Histogram of life expectancy for 145 countries')
```
\newpage

## Goodness of fit test for Totoal Expenditure
```{r}
# Set the sample mean and standard deviation for Total Expenditure
mean_Exp <- mean(data_clean$`Total expenditure`)
sd_Exp <- sd(data_clean$`Total expenditure`)

# Standardized the data of Total expenditure
Exp_std <- (data_clean$`Total expenditure` - mean_Exp) / sd_Exp

# Create 5 equal-probability interval
prob <- c(0:5) * 1/5
qnorm(prob)

# The number of standardized Total expenditure in each interval above
I_1 <- sum(Exp_std <= -0.8416212)
I_2 <- sum(Exp_std > -0.8416212 & Exp_std <= -0.2533471)
I_3 <- sum(Exp_std > -0.2533471 & Exp_std <= 0.2533471)
I_4 <- sum(Exp_std > 0.2533471 & Exp_std <= 0.8416212)
I_5 <- sum(Exp_std > 0.8416212)

# Actual counts
actual_count <- c(I_1, I_2, I_3, I_4, I_5)

# Expected counts
n_row <- nrow(data_clean)
expected <- n_row * 1/5
expected_count <- rep(expected, 5)

# Test statistic
Q <- sum((actual_count - expected_count)^2 / expected_count)

# Calculate P_value
1 - pchisq(Q, df=4)
```
\newpage
## Confidence Interval

CI : $\bar{X}\pm t*\frac{S}{\sqrt{n}}$
```{r}
# Set the sample mean and standard deviation for Total Expenditure
mean_Exp <- mean(data_clean$`Total expenditure`)
sd_Exp <- sd(data_clean$`Total expenditure`)

# Sample size
n <- nrow(data_clean)

# Critical value, here we want to build a 98% confidence interval
t <- qt(c(0.01, 0.99), df = n -1)

# Confidence interval
mean_Exp + t*sd_Exp/sqrt(n)

## Next we divide the dataset data_clean into two groups.

## The first group is all the data whose Total expenditure are lower than 5.569936
group_1 <- data_clean %>%
  filter(`Total expenditure` <= 5.722427) %>%
  select(`Life expectancy`)
summary(group_1)
## The second group is all the data whose Total expenditure are higher than 6.573926
group_2 <- data_clean %>%
  filter(`Total expenditure` >= 6.679562) %>%
  select(`Life expectancy`)
summary(group_2)
```
\newpage
## Two group hypothesis test

test-stats = $\frac{\bar{x_1} - \bar{x_2}}{\sqrt{\frac{(s_1)^2}{n_1} + \frac{(s_2)^2}{n_2}}}$

df = $\frac{(\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2})^2}{\frac{(\frac{s_1^2}{n_1})^2}{n_1-1} + \frac{(\frac{s_2^2}{n_2})^2}{n_2-1}}$

```{r}
# Create a dataset that only contain developed country
developed <- data_clean %>%
  filter(Status == "Developed")

# Create a dataset that only contain developed country
developing <- data_clean %>%
  filter(Status == "Developing")

# Set the mean of Total expenditure for each group
mean_developed <- mean(developed$`Total expenditure`)
mean_developing <- mean(developing$`Total expenditure`)

# Set the standard deviation of total expenditure for each group
sd_developed <- sd(developed$`Total expenditure`)
sd_developing <- sd(developing$`Total expenditure`)

# Set the sample size for each group
n_developed <- nrow(developed)
n_developing <- nrow(developing)

# Test statistic
test_stats <- (mean_developed - mean_developing)/sqrt(sd_developed^2/n_developed 
+ sd_developing^2/n_developing)

# Degrees of freedom
df <- floor((sd_developed^2/n_developed + sd_developing^2/n_developing)^2/
(((sd_developed^2)/n_developed)^2/(n_developed-1)
+ ((sd_developing^2)/n_developing)^2/(n_developing - 1)))

# P-value
1-pt(test_stats, df=df)

summary(developed)
summary(developing)
```
\newpage

## SLR
## (1) Plot

```{r}
# Create a scatter plot to analyze the relationship between Life expectancy and Total expenditure
ggplot(data = data_clean, aes(x = `Life expectancy`, y = `Total expenditure`)) +
  geom_point()+
  labs(x = 'Life expectancy(age)',
       y = 'Expenditure on health of total government expenditure(%)',
       title = 'Relations between Total expenditure and life expectancy')
```
\newpage

## Fitted regression line equation
```{r}
model <- lm(`Life expectancy` ~ `Total expenditure` , data = data_clean)
summary(model)
```

regression line equation: $\hat{y_i} = 69.5846  + 0.9845x_i$
$R^2 = 0.08901$ 

## Hypothesis test for $\beta{_1}$
$H_0: \beta{_1} = 0$
$H_a: \beta{_1} \neq 0$


## Confidence interval for $\beta{_0}$ and $\beta{_1}$
```{r}
# Set the degrees of freedom
df <- model$df.residual

# Set the critical value
t <- qt(c(0.01,0.99), df=df)

# 98% CI for $\beta{_0}$
-60208.5 + t*10950.4

# 98% CI for $\beta{_1}$
981.5 + t* 151.8
```

